---
title: "Map Options"
execute:
  echo: false
  warning: false
format:
  html:
    grid:
      sidebar-width: 0px
      margin-width: 400px
    theme: zephyr
---

```{r setup, include=FALSE}
options(tigris_use_cache = TRUE)
options(bitmapType = "cairo")
library(scales)
library(ptaxsim)
library(gt)
library(DBI)
library(RSQLite)
library(RColorBrewer)
library(patchwork)
library(tigris)
library(tidyverse)
library(glue)
library(sf)

source("map_dist_options.R")
source("../maps.R")
source("../helper_funcs.R")
# set the following variables and run
db_path <- "../ptaxsim-2021.0.4.db"
pin_14 <- "10132040060000"
year <- 2021

ptaxsim_db_conn <- dbConnect(SQLite(), db_path)
```

```{r ptaxsim-data}
pin_details <- lookup_pin(year, pin_14)

render_pin14 <- glue("Parcel ({render_pin(pin_14)})")

pin_exe_details <- pin_details %>%
    select(starts_with("exe")) %>%
    pivot_longer(cols = everything())

pin_exe_total <- pin_exe_details %>%
    pull(value) %>%
    sum()

tax_bill_data <- tax_bill(
    year_vec = year,
    pin_vec = pin_14,
    simplify = F
) %>%
    mutate(agency_total_eav = as.double(agency_total_eav)) %>%
    mutate(agency_major_type = str_to_title(agency_major_type)) %>%
    mutate(tif_agency_name = str_replace_all(tif_agency_name, "TIF - ", "")) %>%
    mutate(tif_agency_name = str_replace_all(tif_agency_name, "RPM1", "Red-Purple Modernisation Phase 1")) %>%
    mutate(tif_agency_name = str_to_title(tif_agency_name)) %>%
    mutate(agency_name = str_to_title(agency_name)) %>%
    filter(agency_total_ext > 0)

cook_county_max_ext <- tax_bill_data %>%
    filter(agency_major_type == "Cook County") %>%
    slice_max(agency_total_ext) %>%
    mutate(
        agency_tax_amount = label_dol_amt(agency_tax_rate * (eav - exe_total)),
        agency_total_eav = label_dol_amt(agency_total_eav),
        agency_total_ext = label_dol_amt(agency_total_ext),
    )

tif_agency_nums <- tax_bill_data %>%
    select(tif_agency_num) %>%
    filter(!is.na(tif_agency_num)) %>%
    unique() %>%
    pull()

```

```{r ptaxsim-geo-data}
cook_roads <- roads("17", "031")

shp_bnd_pin <- get_shp_bnd_pin(pin_14, year)

shp_buf_pin <- shp_bnd_pin %>%
    st_centroid() %>%
    st_transform(3857) %>%
    st_buffer(650) %>%
    st_transform(4326)

shp_bnd_all_schools <- tax_bill_data %>%
    filter(agency_major_type == "School") %>%
    pull(agency_num) %>%
    get_shp_all_schools() %>%
    correct_shp_names() %>%
    mutate(agency_name = as.factor(agency_name))

shp_bnd_tifs <- get_shp_tifs(c(tif_agency_nums, "0")) %>%
    left_join(
        tax_bill_data %>%
            select(tif_agency_num, tif_agency_name) %>%
            filter(!is.na(tif_agency_num)) %>%
            unique(),
        by = join_by(AGENCY == tif_agency_num)
    ) %>%
    mutate(agency_name = str_to_title(tif_agency_name))
```

```{r defaults}
column_colors <- brewer.pal(3, "Pastel1")
column_alpha <- 0.5

map_road_alpha <- 0.3
map_road_lwd <- 0.3
map_fg_fill_alpha <- 0.33
map_bg_fill_alpha <- 0.33
map_color_alpha <- 0.5
map_fg_lwd <- 1.1
map_bg_lwd <- 1.05

map_colors <- c(
    "#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499",
    "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888"
)
map_names <- c(
    render_pin14, "Cook County",
    shp_bnd_all_schools$agency_name %>% as.character(),
    shp_bnd_tifs$agency_name %>% as.character()
)
map_names <- factor(map_names, levels = map_names, ordered = TRUE)
map_colors <- name_map_colors(map_colors, map_names)
css_colors <- paste0(map_colors, "44")
css_colors <- name_map_colors(css_colors, map_names)
```

## Option 1

- If area > 2.5e8 m^2, inset 
- No inset
    - roads + bnd shape + pin shape buffer + pin shape
- Inset
    - Main : roads + bnd shape + pin shape buffer + pin shape + outline
    - Inset: Main + roads + (zoomin)

### Schools

```{r option-1-school}
plt_lst <- list()
for (i in 1:nrow(shp_bnd_all_schools)) {
    plt_lst[[i]] <- map_dist_1(shp_bnd_all_schools[i, ])
}

Reduce("+", plt_lst) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
```

### TIF

```{r option-1-tif}
plt_lst <- list()
for (i in 1:nrow(shp_bnd_tifs)) {
    plt_lst[[i]] <- map_dist_1(shp_bnd_tifs[i, ])
}

Reduce("+", plt_lst) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
```

## Option 2

- Change colors from Option 1 
    - pin shape - black
    - pin shape buffer - no fill; color white

### Schools

```{r option-2-defaults}
map_colors <- c(
    "#333333", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499",
    "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888"
)
map_names <- c(
    render_pin14, "Cook County",
    shp_bnd_all_schools$agency_name %>% as.character(),
    shp_bnd_tifs$agency_name %>% as.character()
)
map_names <- factor(map_names, levels = map_names, ordered = TRUE)
map_colors <- name_map_colors(map_colors, map_names)
css_colors <- paste0(map_colors, "44")
css_colors <- name_map_colors(css_colors, map_names)

```

```{r option-2-school}
plt_lst <- list()
for (i in 1:nrow(shp_bnd_all_schools)) {
    plt_lst[[i]] <- map_dist_2(shp_bnd_all_schools[i, ])
}

Reduce("+", plt_lst) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
```

### TIF

```{r option-2-tif}
plt_lst <- list()
for (i in 1:nrow(shp_bnd_tifs)) {
    plt_lst[[i]] <- map_dist_2(shp_bnd_tifs[i, ])
}

Reduce("+", plt_lst) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
```

## Option 3

- Change colors from Option 1 
    - pin shape - black
    - pin shape buffer - fill alpha 0.35; color white
    - bnd shape - fill alpha 0.5

### Schools

```{r option-3-defaults}
map_road_alpha <- 0.3
map_road_lwd <- 0.3
map_fg_fill_alpha <- 0.5
map_bg_fill_alpha <- 0.5
map_color_alpha <- 0.5
map_fg_lwd <- 1.1
map_bg_lwd <- 1.05

map_colors <- c(
    "#333333", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499",
    "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888"
)
map_names <- c(
    render_pin14, "Cook County",
    shp_bnd_all_schools$agency_name %>% as.character(),
    shp_bnd_tifs$agency_name %>% as.character()
)
map_names <- factor(map_names, levels = map_names, ordered = TRUE)
map_colors <- name_map_colors(map_colors, map_names)
css_colors <- paste0(map_colors, "44")
css_colors <- name_map_colors(css_colors, map_names)

```

```{r option-3-school}
plt_lst <- list()
for (i in 1:nrow(shp_bnd_all_schools)) {
    plt_lst[[i]] <- map_dist_3(shp_bnd_all_schools[i, ])
}

Reduce("+", plt_lst) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
```

### TIF

```{r option-3-tif}
plt_lst <- list()
for (i in 1:nrow(shp_bnd_tifs)) {
    plt_lst[[i]] <- map_dist_3(shp_bnd_tifs[i, ])
}

Reduce("+", plt_lst) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
```


## Option 4

- Change colors from Option 1 
    - pin shape - red
    - pin shape buffer - fill alpha 0.35; color white
    - bnd shape - fill alpha 0.5
- Inset zoom extra

### Schools

```{r option-4-defaults}
map_road_alpha <- 0.3
map_road_lwd <- 0.3
map_fg_fill_alpha <- 0.5
map_bg_fill_alpha <- 0.5
map_color_alpha <- 0.5
map_fg_lwd <- 1.1
map_bg_lwd <- 1.05

map_colors <- c(
    "#C60000", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499",
    "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888"
)
map_names <- c(
    render_pin14, "Cook County",
    shp_bnd_all_schools$agency_name %>% as.character(),
    shp_bnd_tifs$agency_name %>% as.character()
)
map_names <- factor(map_names, levels = map_names, ordered = TRUE)
map_colors <- name_map_colors(map_colors, map_names)
css_colors <- paste0(map_colors, "44")
css_colors <- name_map_colors(css_colors, map_names)

```

```{r option-4-school}
plt_lst <- list()
for (i in 1:nrow(shp_bnd_all_schools)) {
    plt_lst[[i]] <- map_dist_4(shp_bnd_all_schools[i, ])
}

Reduce("+", plt_lst) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
```

### TIF

```{r option-4-tif}
plt_lst <- list()
for (i in 1:nrow(shp_bnd_tifs)) {
    plt_lst[[i]] <- map_dist_4(shp_bnd_tifs[i, ])
}

Reduce("+", plt_lst) +
    plot_layout(guides = "collect") &
    theme(legend.position = "bottom")
```